Comparison of Several Stochastic and Deterministic Derivative-free Global Optimization Algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2019-02-15 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: fix captures of some tables
typography corrected

\begin{document}
%
\title{Comparison of Several Stochastic and Deterministic Derivative-free Global Optimization Algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Vladislav Sovrasov}
%
\authorrunning{V. Sovrasov}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
  \email{sovrasov.vlad@gmail.com}
}%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
In this paper popular open-source solvers are compared against
Globalizer solver, which is developed at the Lobachevsky State University.
The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
competitive performance with other similar solvers. The comparison is done on several sets of
challenging multi-extremal benchmark functions. Also this work considers a method of
heuristic hyperparameters control for the Globalizer allowing to reduce amount of
initial tuning before optimization. The proposed scheme allows substantially increase convergence speed
of the Globalizer by switching between ``local'' and ``global'' search phases in runtime.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{deterministic global optimization $\cdot$ stochastic global optimization
  $\cdot$ algorithms comparison $\cdot$ derivative-free algorithms $\cdot$ black-box
optimization $\cdot$ multi-extremal problems}
\end{abstract}
%
\section{Introduction}

@@

In Ref. \cite{Beiranvand2017}, some general principles were formulated, which, in the author's
opinion, should be obeyed when comparing the optimization methods. In particular, the authors say
about the advantages of the problem generators allowing generating the large sets of problems thus
minimizing the random effects when comparing the methods. At the same time, the use of a single
generator can appear to be not enough for a comprehensive comparison of the methods. In order to
overcome this problem in part, the authors of Ref. \cite{Beiranvand2017} advise to use several
generators of various nature and to create the sets of problems of various complexity.

@@

\begin{table}
\begin{center}
\caption{Class-specific parameters of the optimization algorithms}
  \begin{tabular}{|l|{c}|{c}|{c}|}
    \hline
    & AGS, AGS\(l\) & CRS & DE\\
  \hline
  \(F_{GR}\) & \(r=3\) & popsize=150 & mutation=(1.1,1.9), popsize=60 \\
  \hline
  GKLS 2d Simple & \(r=4.6\) & popsize=200 & mutation=(1.1,1.9), popsize=60 \\
  \hline
  GKLS 2d Hard & \(r=6.5\) & popsize=400 & mutation=(1.1,1.9), popsize=60 \\
  \hline
  GKLS 3d Simple & \(r=3.7\) & popsize=1000 & mutation=(1.1,1.9), popsize=70 \\
  \hline
  GKLS 3d Hard & \(r=4.4\) & popsize=2000 & mutation=(1.1,1.9), popsize=80 \\
  \hline
  GKLS 4d Simple & \(r=4.7\) & popsize=8000 & mutation=(1.1,1.9), popsize=90 \\
  \hline
  GKLS 4d Hard & \(r=4.9\) & popsize=16000 & mutation=(1.1,1.9), popsize=100 \\
  \hline
  GKLS 5d Simple & \(r=4\) & popsize=25000 & mutation=(1.1,1.9), popsize=120 \\
  \hline
  GKLS 5d Hard & \(r=4\) & popsize=30000 & mutation=(1.1,1.9), popsize=140 \\
  \hline
\end{tabular}
  \label{tab:params}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\caption{Number of test optimization problems solved by methods}
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     &  100 & 100 & 76 & 100 & 100 & 97 & 96 & 96 & 67\\
  \hline
  GKLS 2d Simple &  100 & 100 & 85 & 100 & 100 & 100 & 100 & 98 & 90\\
  \hline
  GKLS 2d Hard   &  100 & 97 & 74 & 100 & 100 & 100 & 93 & 85 & 77 \\
  \hline
  GKLS 3d Simple &  100 & 100 & 75 & 100 & 100 & 100 & 89 & 86 & 44 \\
  \hline
  GKLS 3d Hard   &  100 & 100 & 72 & 100 & 99 & 100 & 88 & 77 & 43 \\
  \hline
  GKLS 4d Simple &  100 & 100 & 74 & 100 & 100 & 94 & 82 & 68 & 72 \\
  \hline
  GKLS 4d Hard   &  100 & 100 & 60 & 99 & 99 & 94 & 73 & 55 & 69  \\
  \hline
  GKLS 5d Simple &  100 & 100 & 86 & 100 & 100 & 98 & 100 & 88 & 82  \\
  \hline
  GKLS 5d Hard   &  100 & 100 & 77 & 100 & 93 & 79 & 86 & 77 & 78 \\
  \hline
  \end{tabular}
  \label{tab:solved}
\end{center}
\end{table}

@@

In the present paper, several global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems.
Also a scheme of hyperparameters control for the AGS algorithms was proposed and evaluated.
The results presented in this work allow making the following conclusions:
\begin{itemize}
  \item the proposed modification of the stock AGS, AGS-AR allows to pay less attention to initial hyperparameter tuning and
  performs on-par with properly tuned AGS.
  \item AGS-AR method has demonstrated the convergence
  speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
  implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
\end{itemize}

>>>>

\begin{document}
%
\title{Comparison of Several Stochastic and Deterministic Derivative-free Global Optimization Algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}}
%
\titlerunning{Comparison of Global Optimization Algorithms}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Vladislav Sovrasov}
%
\authorrunning{V. Sovrasov}
%
\institute{Lobachevsky State University of Nizhni Novgorod, Russia \\
  \email{sovrasov.vlad@gmail.com}
}%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
In this paper popular open-source solvers are compared against
Globalizer solver, which is developed at the Lobachevsky State University.
The Globalizer is designed to solve problems with black-box objective function satisfying the Lipschitz condition and shows
competitive performance with other similar solvers. The comparison is done on several sets of
challenging multi-extremal benchmark functions. Also this work considers a method of
heuristic hyperparameters control for the Globalizer allowing to reduce amount of
initial tuning before optimization. The proposed scheme allows substantially increase convergence speed
of the Globalizer by switching between ``local'' and ``global'' search phases in runtime.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{deterministic global optimization $\cdot$ stochastic global optimization
  $\cdot$ algorithms comparison $\cdot$ derivative-free algorithms $\cdot$ black-box
optimization $\cdot$ multi-extremal problems}
\end{abstract}
%
\section{Introduction}

@@

In \cite{Beiranvand2017}, some general principles were formulated, which, in the author's
opinion, should be obeyed when comparing the optimization methods. In particular, the authors say
about the advantages of the problem generators allowing generating the large sets of problems thus
minimizing the random effects when comparing the methods. At the same time, the use of a single
generator can appear to be not enough for a comprehensive comparison of the methods. In order to
overcome this problem in part, the authors of the paper \cite{Beiranvand2017} advise to use several
generators of various nature and to create the sets of problems of various complexity.

@@

\begin{table}
\begin{center}
\caption{Class-specific parameters of the optimization algorithms}
  \begin{tabular}{|l|{c}|{c}|{c}|}
    \hline
    & AGS & CRS & DE\\
  \hline
  \(F_{GR}\) & \(r=3\) & popsize=150 & mutation=(1.1,1.9), popsize=60 \\
  \hline
  GKLS 2d Simple & \(r=4.6\) & popsize=200 & mutation=(1.1,1.9), popsize=60 \\
  \hline
  GKLS 2d Hard & \(r=6.5\) & popsize=400 & mutation=(1.1,1.9), popsize=60 \\
  \hline
  GKLS 3d Simple & \(r=3.7\) & popsize=1000 & mutation=(1.1,1.9), popsize=70 \\
  \hline
  GKLS 3d Hard & \(r=4.4\) & popsize=2000 & mutation=(1.1,1.9), popsize=80 \\
  \hline
  GKLS 4d Simple & \(r=4.7\) & popsize=8000 & mutation=(1.1,1.9), popsize=90 \\
  \hline
  GKLS 4d Hard & \(r=4.9\) & popsize=16000 & mutation=(1.1,1.9), popsize=100 \\
  \hline
  GKLS 5d Simple & \(r=4\) & popsize=25000 & mutation=(1.1,1.9), popsize=120 \\
  \hline
  GKLS 5d Hard & \(r=4\) & popsize=30000 & mutation=(1.1,1.9), popsize=140 \\
  \hline
\end{tabular}
  \label{tab:params}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\caption{Number of test optimization problems solved by the methods}
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     &  100 & 100 & 76 & 100 & 100 & 97 & 96 & 96 & 67\\
  \hline
  GKLS 2d Simple &  100 & 100 & 85 & 100 & 100 & 100 & 100 & 98 & 90\\
  \hline
  GKLS 2d Hard   &  100 & 97 & 74 & 100 & 100 & 100 & 93 & 85 & 77 \\
  \hline
  GKLS 3d Simple &  100 & 100 & 75 & 100 & 100 & 100 & 89 & 86 & 44 \\
  \hline
  GKLS 3d Hard   &  100 & 100 & 72 & 100 & 99 & 100 & 88 & 77 & 43 \\
  \hline
  GKLS 4d Simple &  100 & 100 & 74 & 100 & 100 & 94 & 82 & 68 & 72 \\
  \hline
  GKLS 4d Hard   &  100 & 100 & 60 & 99 & 99 & 94 & 73 & 55 & 69  \\
  \hline
  GKLS 5d Simple &  100 & 100 & 86 & 100 & 100 & 98 & 100 & 88 & 82  \\
  \hline
  GKLS 5d Hard   &  100 & 100 & 77 & 100 & 93 & 79 & 86 & 77 & 78 \\
  \hline
  \end{tabular}
  \label{tab:solved}
\end{center}
\end{table}

@@

In the present paper, several global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems.
Also a scheme of hyperparameters control for the AGS algorithms was proposed and evaluated.
The results presented in this work allow making the following conclusions:
\begin{itemize}
  \item the proposed modification of the stock AGS, AGS-AR allows to pay less attention to initial hyperparameter tuning and
  performs on-par with properly tuned AGS;
  \item AGS-AR method has demonstrated the convergence
  speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
  implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems.
\end{itemize}
